---
title: "Efficient data merging"
author: "Robin Lovelace"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignetted extends Chapter 6 of Efficient R Programming to discuss merging.

The starting point is map data and a mapping package:

```{r, message=FALSE}
library("efficient")
library("dplyr")
library("ggmap")
world = map_data("world")
names(world)
```

Visually compare this new dataset of the `world` with `ghg_ems` (e.g. via `View(world); View(ghg_ems)`). It is clear that the column `region` in the former contains the same information as `Country` in the latter. This will be the *joining variable*; renaming it in `world` will make the join more efficient. 

```{r, echo=-1}
data(ghg_ems, package = "efficient")
world = rename(world, Country = region)
ghg_ems$All = rowSums(ghg_ems[3:7])
```

```{block same-class, type = "rmdtip"}
Ensure that both joining variables have the same class (combining `character` and `factor` columns can cause havoc).
```

How large is the overlap between `ghg_ems$Country` and `world$Country`? We can find out using the `%in%` operator, which finds out how many elements in one vector match those in another vector. Specifically, we will find out how many *unique* country names from `ghg_ems` are present in the `world` dataset:

```{r}
unique_countries_ghg_ems = unique(ghg_ems$Country)
unique_countries_world = unique(world$Country)
matched = unique_countries_ghg_ems %in% unique_countries_world
table(matched)
```

This comparison exercise has been fruitful: most of the countries in the `co2` dataset exist in the `world` dataset. But what about the 20 country names that do not match? We can identify these as follows:

```{r}
(unmatched_countries_ghg_ems <- unique_countries_ghg_ems[!matched])
```

It is clear from the output that some of the non-matches (e.g. the European Union) are not countries at all. However, others, such as 'Gambia, The' and the United States clearly should have matches. *Fuzzy matching* can help find which countries *do* match, as illustrated the first non-matching country below:

```{r}
(unmatched_country = unmatched_countries_ghg_ems[1])
unmatched_world_selection = agrep(pattern = unmatched_country, unique_countries_world, max.distance = 10)
unmatched_world_countries = unique_countries_world[unmatched_world_selection]
```

What just happened? We verified that first unmatching country in the `ghg_ems` dataset was not in the `world` country names. So we used the more powerful `agrep` to search for fuzzy matches (with the `max.distance` argument set to `10`. The results show that the country `Antigua & Barbuda` from the `ghg_ems` data matches *two* countries in the `world` dataset. We can update the names in the dataset we are joining to accordingly:

```{r}
world$Country[world$Country %in% unmatched_world_countries] =
  unmatched_countries_ghg_ems[1]
```

The above code reduces the number of country names in the `world` dataset by replacing *both* "Antigua" and "Barbuda" to "Antigua & Barbuda". This would not work other way around: how would one know whether to change "Antigua & Barbuda" to "Antigua" or to "Barbuda".

Thus fuzzy matching is still a laborious process that must be complemented by human judgement. It takes a human to know for sure that `United States` is represented as `USA` in the `world` dataset, without risking false matches via `agrep`. 

To fix the remaining issues, we simply continued with the same method, using a `for` loop and verifying the results instead of doing all by hand. The code used to match the remaining unmatched countries can be seen on the book's [GitHub page](https://github.com/csgillespie/efficientR/blob/master/05-data-carpentry.Rmd).

```{r co2-clean, echo=FALSE, eval=FALSE}
i = n[1]
match_df = tibble(co2_name = n, w_name = NA)
for(i in n){
  (fm = agrep(i, w_u, max.distance = 10))
  (w_um = w_u[fm])
  match_df$w_name[match_df$co2_name == i] = paste(w_um, collapse = "|")
  # world$Country[world$Country %in% w_u1] = i
}

# View(match_df) # check the results: 1, 3 , 14, 16, 17 are right
i = 3
for(i in c(1, 3 , 14, 16, 17)){
  world$Country[grep(match_df$w_name[i], world$Country) ] =
    match_df$co2_name[i]
}
match_df = match_df[-c(1, 3 , 14, 16, 17),]

# manually fix countries with multiple matches
world$Country[grep("Baham", world$Country)] =
  c_u[grep("Baham", c_u)]
world$Country[grep("Democratic Republic of the Congo", world$Country)] =
  c_u[grep("Congo, Dem. Rep.", c_u)]
world$Country[grep("Republic of C", world$Country)] =
  c_u[grep("Congo, R", c_u)]
world$Country[grep("Ivo", world$Country)] =
  c_u[grep("Ivo", c_u)]
world$Country[grep("Gambia", world$Country)] =
  c_u[grep("Gambia", c_u)]
world$Country[grep("Macedonia", world$Country)] =
  c_u[grep("Macedonia", c_u)]
world$Country[grep("USA", world$Country)] =
  c_u[grep("United States", c_u)]
world$Country[grep("UK", world$Country)] =
  c_u[grep("United Kingdom", c_u)]
world$Country[grep("North Korea", world$Country)] =
  c_u[grep("Korea, Dem. Rep. \\(N", c_u)]
world$Country[grep("South Korea", world$Country)] =
  c_u[grep("Korea, Rep", c_u)]
world$Country[grep("Russia", world$Country)] =
  c_u[grep("Russia", c_u)]
world$Country[grep("Vincent", world$Country)] =
  c_u[grep("Vincent", c_u)]

# ghg_ems = ghg_ems[!ghg_ems$Country == "World",]

# save the result as 'm', for match
c_u = unique(ghg_ems$Country)
w_u = unique(world$Country)
m = c_u %in% w_u
# summary(m)
n = c_u[!m]
```


There is one more stage that is needed before global CO^2^ emissions can be mapped for any year: the data must be *joined*. The base function `merge` can do this but we strongly recommend using one of the `join` functions from **dplyr**, such as `left_join` (which keeps all rows in the original dataset) and `inner_join` (which keeps only rows with matches in both datasets). This is a very clear case of when **dplyr** is advantageous: `merge()`'s interface is complicated, the code is less readable, and `*_join` functions are faster. `inner_` (which keeps all rows in both datasets for which there are matches) and `left_` (which keeps only) `join()` methods are illustrated below:

```{r}
nrow(world)
nrow({world_co2 = left_join(world, ghg_ems)})
nrow(inner_join(world, ghg_ems))
```

Note that `inner_join` removes rows from the `world` dataset which have no match in `ghg_ems`: if we were to plot the resulting dataset, the continent of Antarctica and a number of countries not represented in the `ghg_ems` dataset would be absent. Figure \@ref(fig:6-1) shows the results of this data carpentry, produced using a modified version of the **ggplot2** code below, were worth the effort.

```{r, eval=FALSE}
world_co2_2012 = filter(world_co2, Year == 2012 | is.na(Year))
ggplot(world_co2_2012, aes(long, lat)) +
  geom_polygon(aes(fill = All, group = group))
```

```{r, echo=FALSE}
library("scales")
world_co2_2012 = filter(world_co2, Year == 2012 | is.na(Year))
ggplot(world_co2_2012, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = All)) +
  geom_path(size = 0.2) +
  scale_fill_gradient(
    low = "blue", high = "red", trans = "log",
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    name = expression(MtCO[2])) +
  coord_equal() +
  theme_nothing(legend = TRUE)
# ggsave("figures/world_co2.png")
```