---
title: "Answers to exercises in Efficient R Programming: Chapter 6"
author: "Colin Gillespie and Robin Lovelace"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Chapter 6 solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

#### Exercise {-}

Create the following data frame

```{r, eval=FALSE}
df_base = data.frame(colA = "A")
```

Try and guess the output of the following commands

```{r results="hide", eval=FALSE}
print(df_base)
df_base$colA
df_base$col
df_base$colB
```

Now create a tibble data frame and repeat the above commands.

```{r, eval=FALSE}
library("tibble")
df_tibble = tibble::tibble(colA = "A")
```
    
The tibble does not guess the column names:
    
```{r results="hide", eval=FALSE}
print(df_tibble)
## # A tibble: 1 × 1
##    colA
##   <chr>
## 1     A
df_tibble$colA
## [1] "A"
df_tibble$col
## Warning: Unknown column 'col'
df_tibble$colB
## Warning: Unknown column 'colB'
```

#### Exercises {-}

1. What are the three criteria of tidy data?

    Each variable forms a column
    
    Each observation forms a row
    
    Each unit of analysis (e.g. person, household) forms a single table
    
    These answers can be found in the **tidyr** vignette:

    ```{r, eval=FALSE}
    vignette(package = "tidyr")
    vignette("tidy-data")
    ```

2. Load and look at subsets of these datasets. The first is the `pew` datasets we've been using already. The second reports the points that define, roughly, the geographical boundaries of different London boroughs. What is 'untidy' about each?

```{r, eval=FALSE}
data(pew, package = "efficient")
head(pew, 10)
data(lnd_geo_df, package = "efficient")
head(lnd_geo_df, 10)

##                    name_date population        x        y
## 1               Bromley-2001     295535 544361.6 172379.2
## 2               Bromley-2001     295535 549546.0 169911.0
## 3               Bromley-2001     295535 539596.3 160796.3
## 4               Bromley-2001     295535 533693.2 170730.1
## 5               Bromley-2001     295535 533717.7 170813.8
## 6               Bromley-2001     295535 534003.7 171442.1
## 7               Bromley-2001     295535 541104.7 173355.7
## 8               Bromley-2001     295535 544361.6 172379.2
## 9  Richmond upon Thames-2001     172330 523604.7 176320.7
## 10 Richmond upon Thames-2001     172330 521454.5 172361.5
```

In the `pew` dataset, not every column is a variable: the single variable of income band is represented by many columns.

3. Convert each of the above datasets into tidy form.

    As illustrated previously, `gather()` converts wide data into long, tidy forms:
    
    
```{r, eval=FALSE}
library("tidyr")
gather(pew, Income, count, -religion) # results not shot shown
```

In the `lnd_geo_df` dataset, not every column is a distinct variable: `name_date` contains two variables (the name of the zone and the date). This can be fixed as follows:
    
```{r, eval=FALSE}
tidy_geo = separate(lnd_geo_df, name_date,
                  into = c("name", "date"), sep = -5)
unique(tidy_geo[1:2])[1:3,]

##                     name date
## 1               Bromley- 2001
## 9  Richmond upon Thames- 2001
## 15           Hillingdon- 2001
```

4. Consider the following string of phone numbers and fruits [from the **stringr** package](http://stringr.tidyverse.org/reference/str_match.html):

```{r}
strings = c(" 219 733 8965", "329-293-8753 ", "banana", "595 794 7569",
             "387 287 6718", "apple", "233.398.9187  ", "482 952 3315", "239 923 8115",
             "842 566 4692", "Work: 579-499-7527", "$1000", "Home: 543.355.3679")
```

Write functions in **stringr** and base R that return:

- A logical vector reporting whether or not each string contains a number.
- Complete words only, without extraneous non-letter characters.

    ```{r, eval=FALSE}
    library(stringr)
    str_detect(string = strings, pattern = "[0-9]")
    str_extract(strings, pattern = "[A-z]+")
    ```

1. Try running each of the chaining examples above line-by-line, so the first two entries for the first example would look like this:

```{r, eval=FALSE}
library("dplyr")
data("wb_ineq", package="efficient")
```

followed by:

```{r, eval=FALSE}
wb_ineq %>% 
  select(Year, gini) 
```

Explain in your own words what changes each time.

```{r, eval=FALSE}
wb_ineq %>% 
  select(Year, gini) %>% # select only named columns
  mutate(decade = floor(Year / 10) * 10) %>% # create decade column
  group_by(decade) %>% # group the dataset
  summarise(mean(gini, na.rm = TRUE)) # find the mean gini index per year

## # A tibble: 6 × 2
##   decade `mean(gini, na.rm = TRUE)`
##    <dbl>                      <dbl>
## 1   1970                   40.05333
## 2   1980                   37.84583
## 3   1990                   42.02168
## 4   2000                   40.54188
## 5   2010                   39.28561
## 6     NA                        NaN
```

2. Use chained **dplyr** functions to answer the following question: In which year did countries without and 'a' in their name have the lowest level of inequality?

    1981, as illustrated below:

```{r, eval=FALSE}
wb_ineq %>% 
  filter(!grepl("a", Country)) %>%
  filter(!is.na(gini)) %>% 
  group_by(Year) %>% 
  summarise(mean_gini = min(mean(gini), na.rm = TRUE)) %>% 
  top_n(-1, mean_gini)

## # A tibble: 1 × 2
##    Year mean_gini
##   <int>     <dbl>
## 1  1981     23.29
```

    Note: the question is open to interpretation. In the above answer, 'lowest level of inequality' is interpreted as 'mean gini index for all countries without a in their name per year'.

#### Exercises {-}

1. Refer back to the greenhouse gas emissions example, in which we found the top 3 countries in terms of emissions growth in the transport sector. a) Explain in words what is going on in each line. b) Try to find the top 3 countries in terms of emissions in 2012 - how is the list different?

    Russia is now present on the list - it has high absolute transport emissions, but relatively low levels of emissions growth.

```{r, message=FALSE}
library("dplyr")
data("ghg_ems", package = "efficient")
ghg_ems %>%
  filter(!grepl("World|Europe", Country)) %>% 
  filter(Year == 2012) %>% 
  top_n(3, Transportation) %>%
  arrange(desc(Transportation))

## # A tibble: 3 × 7
##              Country  Year Electricity Manufacturing Transportation  Other
##                <chr> <dbl>       <dbl>         <dbl>          <dbl>  <dbl>
## 1      United States  2012     2369.60        495.41        1667.26 541.88
## 2              China  2012     4404.92       2546.06         702.91 551.97
## 3 Russian Federation  2012      995.08        293.48         235.22 135.25
## # ... with 1 more variables: Fugitive <dbl>
```

2. Explore **dplyr**'s documentation, starting with the introductory vignette, accessed by entering [`vignette("introduction")`](https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html).

3. Test additional **dplyr** 'verbs' on the `wb_ineq` dataset. (More vignette names can be discovered by typing `vignette(package = "dplyr")`.)

#### Exercise {-}

- Follow the worked example below to create and query a database on land prices in the UK using **dplyr** as a front end to an SQLite database. 
The first stage is to read-in the data:

```{r, eval=FALSE}
# See help("land_df", package="efficient") for details
data(land_df, package="efficient")
```

The next stage is to create an SQLite database to hold the data:

```{r, eval=FALSE}
# install.packages("RSQLite") # Requires RSQLite package
my_db = src_sqlite("land.sqlite3", create = TRUE)
land_sqlite = copy_to(my_db, land_df, indexes = list("postcode", "price")) 
```

What class is the new object `land_sqlite`?

    It's a table with additional attributes to show it's also an SQLite database.

```{r, eval=FALSE}
class(land_sqlite)
```


Why did we use the `indexes` argument?

    For fast sorting on these variables.

From the above code we can see that we have created a `tbl`. This can be accessed using **dplyr** in the same way as any data frame can. Now we can query the data. You can use SQL code to query the database directly or use standard **dplyr** verbs on the table.

```{r, eval=FALSE}
# Method 1: using sql
tbl(my_db, sql('SELECT "price", "postcode", "old/new"  FROM land_df'))
```

How would you erform the same query using `select()`? Try it to see if you get the same result (hint: use backticks for the `old/new` variable name).

```{r, eval=FALSE}
# Method 2: using dplyr
select(land_sqlite, price, postcode, `old/new`)
```

