---
title: "Answers to exercises in Efficient R Programming"
author: "Colin Gillespie and Robin Lovelace"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Chapter 8 solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Chapter 8 solutions}
-->

### Exercises {-}

The following two exercises aim to help you determine if it is worthwhile upgrading your RAM.

1. R loads everything into memory, i.e. your computers RAM. How much RAM does your computer you have?

    The **benchmarkme** package provides a function for obtaining the amount of RAM on your system:
    ```{r results="hide"}
    benchmarkme::get_ram()
    ```

2. Using your preferred search engine, how much does it cost to double the amount of available RAM on your system? 

    16GB of RAM costs around £90/$110.


### Exercises {-}

These exercises aim to condense the previous section into the key points.

1. Are you using $32$-bit or $64$-bit version of R?

    An easy of to tell is to look at `.Machine$sizeof.pointer`:
    ```{r}
    if(.Machine$sizeof.pointer == 8) 64 else 32
    ```
2. If you are using Windows, what are the results of running the command `memory.limit()`?


### Exercise {-}

To assess whether you should consider cloud computing, how much does it cost to rent a machine comparable to your laptop in the cloud?

  * [On-demand](https://aws.amazon.com/ec2/pricing/on-demand/) amazon service. 
  * Spot instances are much cheaper; you bid for compute power.

