---
title: "Answers to exercises in Efficient R Programming"
author: "Colin Gillespie and Robin Lovelace"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Chapter 2 solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Chapter 2 solutions}
-->


Welcome to the Efficient R Programming answers sheet.
Here you will find answers to the numerous exercises provided in the book Efficient R Programming.

## Chapter 2

### R Version

- What version of R are you using? Is it the most up-to-date?

Your R version is stored in a variable called `R.version`. Enter these commands to find out the version.

```{r}
R.version
R.version.string
```

This can be compared with the latest version, which is described on R's website: [cran.r-project.org/](https://cran.r-project.org/). If you've not got the latest version, consider installing the latest version!

- Do any of your packages need updating?

To find out which (if any) of your packages need updating use the following command:

```{r, eval=FALSE}
old.packages()
```

### R Startup

- What are the three locations where the startup files are stored? Where are these locations on your computer?

The three locations where startup files can be stored and used are `R_HOME`, `HOME` and the project's working directory. The specific locations of these places will vary from one computer to another but can be found using the following commands:

```{r}
R.home()
path.expand("~")
getwd()
```

See ?Startup for more details


- For each location, does a `.Rprofile` or `.Renviron` file exist?

Navigate to the locations outlined above to find out.

- Create a `.Rprofile` file in your current working directory that prints the message `Happy efficient R programming` each time you start R at this location.

To test if this works, restart the session (e.g. via `Shift+Ctrl+F10` in RStudio) and see if the message appears. Creating the following line in a file called `.Rprofile` in your project's root directory will do this:

```{r}
message("Happy efficient R programming")
```

- What happens to the startup files in `R_HOME` if you create them in `HOME` or local project directories? 

The `R_HOME` startup files will be ignored. Likewise the `HOME` startup files will be ignored if there are local startup files.

### RStudio

The first set of six exercises is about exploring and trying out RStudio interactively. There is no right answer in this case - the best approach is to play around. 

- Try modifying the look and appearance of your RStudio setup.

Again, there is no right answer to this question. It's worth experimenting a little until the setup is right for you. But it's not worth spending all day doing that from an efficiency perspective!

- What is the keyboard shortcut to show the other shortcut? (Hint: it begins with `Alt+Shift` on Linux and Windows.)

On Linux and Windows machines `Alt+Shift+K` will show commonly used shortcuts.

- Try as many of the shortcuts revealed by the previous step as you like. Write down the ones that you think will save you time, perhaps on a post-it note to go on your computer.

Again this is highly subjective and only a few handy examples are mentioned here. `Ctl+Up` can be particularly useful for navigating previous commands that have been partially entered. `Ctl+D` deletes the current line while `Alt+Shift+Down` copies it down.

### BLAS 

### Exercises

- What BLAS system is your version of R using?

You can find out using the **benchmarkme** package:

```{r}
benchmarkme::get_linear_algebra()
```

