---
title: "Answers to exercises in Efficient R Programming"
author: "Colin Gillespie and Robin Lovelace"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Chapter 3 solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Chapter 3 solutions}
-->
```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE, 
  fig.align="center",
  fig.pos="t"
)
library(microbenchmark)
```

#### Exercise {-}

Use the **microbenchmark** package to compare the vectorised construct `x = x + 1`, to the
`for` loop version. Try varying the size of the input vector.

```{r}
add_vec = function(x) x + 1
add_for = function(x) {
  for(i in seq_along(x))
    x[i] = x[i] + 1
  x
}
x = rnorm(10000)
microbenchmark(add_vec(x), add_for(x))
```


#### Exercises {-}

1. Time the two methods for calculating the log sum. 
1. What happens when the `length(x) = 0`, i.e. we have an empty vector?

    ```{r}
    x = c()
    log(x[0])
    ```

### Exercise {-}

Verify that `monte_carlo_vec()` is faster than `monte_carlo()`. How does this relate to 
the number of darts, i.e. the size of `N`, that is used


### Exercises {-}

The `stop()` function has an argument `call.` that indicates if the function call
should be part of the error message. Create a function and experiment with this option.


#### Exercise {-}

Factors are slightly more space efficient than characters. Create a character vector
and corresponding factor and use `pryr::object_size()` to calculate the space needed for
each object.

```{r echo=FALSE, eval=FALSE}
ch = sample(month.name, 1e6, replace = TRUE)
fac = factor(ch, levels = month.name)
pryr::object_size(ch)
pryr::object_size(fac)
```

#### Exercises {-}

1. Rewrite the `sapply()` function calls above using `vapply()` to ensure type consistency.

1. How would you make subsetting data frames with `[` type consistent? Hint: look at
the `drop` argument.

#### Exercise {-}

Construct a box plot of timings for the standard plotting function and the memoised
version. 

#### Exercise {-}

1. Write a stop-watch function __without__ using function closures.
1. Many stop-watches have the ability to measure not only your overall time but also you
individual laps. Add a `lap()` function to the `stop_watch()` function that will record
individual times, while still keeping track of the overall time.
