---
title: "Answers to exercises in Efficient R Programming"
author: "Colin Gillespie and Robin Lovelace"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Chapter 1 solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Chapter 1 solutions}
-->

#### Exercises

Consider the following benchmark to evaluate different functions for calculating the cumulative sum of the whole numbers from 1 to 100:

```{r}
x = 1:100 # initiate vector to cumulatively sum

# Method 1: with a for loop (10 lines)
cs_for = function(x){
  for(i in x){
    if(i == 1){
      xc = x[i]
    } else {
      xc = c(xc, sum(x[1:i]))
    }
  }
  xc
}

# Method 2: with apply (3 lines)
cs_apply = function(x){
  sapply(x, function(x) sum(1:x))
}

# Method 3: cumsum (1 line, not shown)
microbenchmark::microbenchmark(cs_for(x), cs_apply(x), cumsum(x))
```

1. Which method is fastest and how many times faster is it?

2. Run the same benchmark, but with the results reported in seconds, on a vector of all the whole numbers from 1 to 50,000. Hint: also use the argument `neval = 1` so that each command is only run once to ensure the results complete (even with a single evaluation the benchmark may take up to or more than a minute to complete, depending on your system). Does the *relative* time difference increase or decrease? By how much?

```{r, eval=FALSE, echo=FALSE}
x = 1:5e4 # initiate vector to cumulatively sum
microbenchmark(cs_for(x), cs_apply(x), cumsum(x), times = 1, unit = "s")
```

3. Test how long the different methods for subsetting the data frame `df`, presented in Section \@ref(benchmarking-example), take on your computer. Is is faster or slower at subsetting than the computer on which this book was compiled?

4. Use `system.time()` and a `for()` loop to test how long it takes to perform the subsetting operation 50,000 times. Before testing this, do you think it will be more or less than 1 second, for each subsetting method? Hint: the test for the first method is shown below:

```{r, eval=FALSE}
# Test how long it takes to subset the data frame 50,000 times:
system.time(
  for(i in 1:50000){
    df[3, 2]
  }
)
```

5. Bonus exercise: try profiling a section of code you have written using **profvis**. Where are the bottlenecks? Were they where you expected?